---
title: "SSC442_Lab02"
author: "Dan Laird"
date: "1/29/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
library(dplyr)


ameslist <- read.table("https://msudataanalytics.github.io/SSC442/Labs/data/ames.csv",
                       header = TRUE,
                       sep = ",")
Ames <- select(ameslist,c(Id, LotFrontage, LotArea, OverallQual, OverallCond, YearBuilt, YearRemodAdd, 
  BsmtFinSF1, BsmtUnfSF, TotalBsmtSF,X1stFlrSF, X2ndFlrSF, BsmtFullBath, BsmtHalfBath,
  FullBath, HalfBath, BedroomAbvGr, KitchenAbvGr, TotRmsAbvGrd,Fireplaces, GarageYrBlt, GarageCars,
  GarageArea, WoodDeckSF, OpenPorchSF, EnclosedPorch, ScreenPorch, PoolArea, MoSold, YrSold,
  SalePrice))

#Subset data for matrix
Pricecorr <- select(Ames,c(LotArea,OverallQual,BsmtFinSF1,GarageArea,WoodDeckSF,Fireplaces,
              GarageCars,PoolArea, FullBath, OpenPorchSF, TotRmsAbvGrd, SalePrice))

#correlation matrix
pairs(Pricecorr)
```


```{r, echo=FALSE}
pairs(Pricecorr)
```
Correlation Matrix


```{r, echo=FALSE}
cor(Pricecorr)
```

For these results, I would have expected pool area and wood dech square footage to be stronger predictors of sale price, but as it turns out, some of the strongest predictors are the overall quality measure, and garage area.  Total rooms above ground also shows a decently strong correlation, but things like number of bathrooms and Finished basement square footage had less of an effect than expected on sale price.  It is surprising to see information relating to the garage being some of the strongest predictors based on house characteristics.


```{r, echo=FALSE}
plot(ameslist$SalePrice, ameslist$GrLivArea, main="Price vs GR Living Area",
     xlab="Sale Price ", ylab="Gr Living Area ", pch=19)
abline(abline(lm(GrLivArea~SalePrice, data=ameslist),col="red"))

```
The largest outlier is a house with over 5000 GR Living area, but sold for less than 200,000 dollars. Upon further inspection, It is a 2 story house, with large square footage, an unfinished basememt, large garage, and a pool.  In all, it seems like a very nice house, but it was sold in 2008, so there is a good chance that the recession is the reason this house sold for so little.
```{r, echo=FALSE}
lm(SalePrice ~ GarageType, data=ameslist)
```
Coefficients:
  (Intercept)   GarageTypeAttchd  GarageTypeBasment  GarageTypeBuiltIn  GarageTypeCarPort  
151283              51609               9287             103468             -41321  
GarageTypeDetchd  
-17192  
```{r, echo=FALSE}
multreg <- lm(SalePrice ~ (LotFrontage+ LotArea+ OverallQual+OverallCond+ YearBuilt+ YearRemodAdd+
                                BsmtFinSF1+ BsmtUnfSF+ TotalBsmtSF+X1stFlrSF+ X2ndFlrSF+ BsmtFullBath+BsmtHalfBath+
                                FullBath+ HalfBath+ BedroomAbvGr+ KitchenAbvGr+ TotRmsAbvGrd+Fireplaces+ GarageYrBlt+ GarageCars+
                                GarageArea+ WoodDeckSF+ OpenPorchSF+ EnclosedPorch+ ScreenPorch+ PoolArea+ MoSold+ YrSold+
                                SalePrice), data=ameslist)


summary(multreg)
```
    It looks like the strongest relationships between sale price and these variables 
exist among factors like Square footage, how many cars the garage holds, the lot area, and 
the total number of rooms above ground.  The year variable suggests that for every year
after being built, houses will depreciate in value by the smount of the coefficient 
(in this model, at least.)
```{r, echo=FALSE}
plot(multreg)
```
The fit is mostly good until you start to reach x values of 400000 and over.
 These points would be outliers if the fit of the model was more accurate, but in this case,
all of the values take a sharp upward turn un this area.  This model leaves out these higher valued estimates.  There are some outliers with low residual values that could be skewing the fit in this direction, so this may explain the problem.
```{r, echo=FALSE}
multregint <- lm(SalePrice ~ (LotFrontage+ LotArea+ OverallQual+OverallCond+ YearBuilt+ YearRemodAdd+
                             BsmtFinSF1+ BsmtUnfSF+ TotalBsmtSF+X1stFlrSF* X2ndFlrSF+ BsmtFullBath+BsmtHalfBath+
                             FullBath* HalfBath+ BedroomAbvGr+ KitchenAbvGr+ TotRmsAbvGrd+Fireplaces+ GarageYrBlt+ GarageCars+
                             GarageArea+ WoodDeckSF+ OpenPorchSF+ EnclosedPorch+ ScreenPorch+ PoolArea+ MoSold+ YrSold+
                             SalePrice), data=ameslist)


plot(multregint)
```
It may make sense to add interaction terms between variables like 1st floor square footage and 2nd floor square footage, given that second floor SF depends on the existance of a 1st floor.  Also, half-baths and full-baths may be able to be modeled with an interaction term involved, because very few houses just have one or the other.  Overall, these terms don't help out model a lot, but it is worth exploring some of these factors.
```{r, echo=FALSE}
multreg_2 <- lm(log(SalePrice) ~ (LotFrontage+ LotArea+ OverallQual+OverallCond+ YearBuilt+ YearRemodAdd+
                                    BsmtFinSF1+ BsmtUnfSF+ TotalBsmtSF+X1stFlrSF+ X2ndFlrSF+ BsmtFullBath+BsmtHalfBath+
                                    FullBath+ HalfBath+ BedroomAbvGr+ KitchenAbvGr+ TotRmsAbvGrd+Fireplaces+ GarageYrBlt+ GarageCars+
                                    GarageArea+ WoodDeckSF+ OpenPorchSF+ EnclosedPorch+ ScreenPorch+ PoolArea+ MoSold+ YrSold+
                                    SalePrice), data=ameslist)

plot(multreg_2)

```
Taking the natural log of SalePrice gives a better fit to the residuals. This especially helps with residuals since the data is skewed right.

